The semantic drift of quotations in blogspace: a case study in short-term cultural evolution

### 2016-07-30 brains-introduction.tex
passed over introduction, related work, methods

\tb{The reciprocal influence between cognition and culture has a long history in both social science and psychology. While this question has been the subject of intense debate in the social sciences in the 20th century (starting with Durkheim's initial works, \citeyear{durkheim_les_1912}, later tackled in earnest by \hbox{e.g.} Mauss' \emph{Techniques of the Body}, \citeyear{mauss_les_1936}, Giddens' \emph{Structuration Theory}, \citeyear{giddens_constitution_1984}, and Bourdieu's \emph{Sens Pratique}, \citeyear{bourdieu_sens_1980}), today's discussion is mostly structured by proponents from cognitive science.}
%Since the very beginnings of both social science and psychology, scientists have tried to capture the way cognition and culture influence each other. The question has been the subject of intense debate in the social sciences in the 20th century (starting with Durkheim's initial works, \citeyear{durkheim_les_1912}, later tackled in earnest by \hbox{e.g.} Mauss' \emph{Techniques of the Body}, \citeyear{mauss_les_1936}, Giddens' \emph{Structuration Theory}, \citeyear{giddens_constitution_1984}, and Bourdieu's \emph{Sens Pratique}, \citeyear{bourdieu_sens_1980}), but today's discussion is mostly structured by proponents from cognitive science.
These construe culture as an evolutionary process analogous and parallel to biological evolution \tbn{(and especially the modern synthesis' account of it)}.
That analogy can \tb{also} be traced a long way back \tbn{in the 20th century and earlier}, with milestones such as Kroeber's works~\citeyearpar{kroeber_nature_1952}, Dawkins' \emph{Memetics}~\tb{\citeyearpar{dawkins_selfish_1976}}, and later the development of \emph{Dual Inheritance Theory} by \citet{boyd_culture_1985} and \citet{cavalli-sforza_cultural_1981} among others.
More recently, Dan Sperber has drawn on this principle to explicitly connect anthropology and cognitive science through the theory of \emph{Epidemiology of Representations} \citep{sperber_explaining_1996}, and the study of cultural evolution has been growing steadily since.

The collection of works by \citet{aunger_darwinizing_2000} (in particular \citealp{bloch_well-disposed_2000}, and \citealp{kuper_if_2000}) has shown how the theory of memetics cannot account for the levels of transformation culture undergoes as it is transmitted.
\Citet{mesoudi_multiple_2008} have discussed the uses of transmission chain experiments to test what dual inheritance theory can explain about cultural evolution.
\Citet{morin_how_2013} and \citet{miton_universal_2015}, by carefully compiling a series of anthropological works, show how cognitive biases have influenced the evolution of cultural artifacts over several centuries.
\citeauthor{kirby_cumulative_2008}~(\citeyear{kirby_cumulative_2008,cornish_systems_2013}) have shown how evolutionary pressures lead to the emergence of structured and expressive artificial languages in simulations and laboratory experiments.
Such transmission chain experiments have also been explored in non-human primates by \citet{claidiere_cultural_2014}.

The theory of epidemiology of representations proposes a unifying framework for all these works by recasting them as questions of spread and transformation of representations:
these are alternatively located in the mind (``mental representations'' in Sperber's terminology), or in the outer world (``public representations'') as expressions of mental representations in diverse cultural artifacts (pieces of text, utterances, pictures, building techniques, etc.).
A human society is then modeled as a large dynamical system of people constantly interpreting public representations into mental representations, and producing new public representations based on what they have previously interpreted.
Two key points are that (a) transmission is not reliable (representations change significantly each time they are interpreted and produced anew, as opposed to \hbox{e.g.} memetics), and (b) the reciprocal influences of cognition and culture can be captured by studying the evolution of public representations themselves, which is what the studies cited above are doing.

The theory makes an additional strong hypothesis, which this paper focuses on:
as transformations accumulate, some representations evolve to be very stable and spread throughout an entire society without changing any more (they are called ``cultural representations'', because they characterize a given culture).
This process should manifest itself as attractors (called ``cultural attractors'') in the dynamical system that models cultural evolution, that is:
there should be areas of the representation space where cognitive effects in transformations bring representations closer to a given stable asymptotic point.\footnote{
\newtext{Attractors need not be points in fact, they can also be sub-areas; in that case any transformation brings representations in the area closer to (or maintained inside) the target sub-area.
}}

This hypothesis, a cornerstone of the theory because of the intelligibility it gives to cultural evolution, has been hard to test in concrete situations as quantitative data on out-of-laboratory cultural artifacts is not easy to collect.
One approach, as mentioned above, has been the meta-analysis of large bodies of anthropological studies (see \citealp{miton_universal_2015}, for instance).
This paper exemplifies a second approach, taking advantage of the ever-increasing avalanche of available digital footprints since the 2000's.
Indeed, tools and computing power to analyze such data are now widespread, and the body of research aimed at describing online communities and content is growing accordingly.
For instance, the propagation of cultural artifacts across social networks has been studied in blogspace~\citep{gruhl_information_2004} and in the email network~\citep{liben-nowell_tracing_2008};
\citet{cointet_socio-semantic_2009} have described the reciprocal influence between the social network topology and the distribution of issues;
\citet{leskovec_meme-tracking_2009} detailed the characteristic times and diffusion cycles both within these social networks and with respect to the topical dynamics of news media, and \citet{danescu-niculescu-mizil_you_2012} have studied the characteristics of particularly memorable quotes that circulate in those networks.
We believe those works can connect the field of cultural evolution with psycholinguistics to advance the testing of cultural attractors.

@@

To show this we analyze the way quotes in blogs and media outlets are modified when they are copied from website to website.
These public representations should normally not change as they spread on the Web (as opposed to more elaborate expressions or opinions, not identified as quoted utterances), but empirical observation shows that they are in fact occasionally transformed~\citep{simmons_memes_2011}:
authors spontaneously transform quotes, not only cropping them but also replacing words.
For instance the quote ``we will not be scared of these cowards'' (a substring of a quote from former Pakistani President Asif Ali Zardari) is also found as ``we will not be \textbf{afraid} of these cowards''.
More meaningful changes often happen too, such as the transformation of McCain's ``I admire Senator Obama and his accomplishments'' during the 2008 U.S. presidential campaign, into ``I \textbf{respect} Senator Obama and his accomplishments''.
Since authors are implicitly required to copy quotes exactly, we can assume that most transformations, especially simple ones such as those shown above, are the result of automatic (\hbox{i.e.} hard to control) low-level cognitive biases of the authors.

Our question is as follows: given such representations that seem to evolve precisely because of the kind of automatic cognitive biases referred to in the theory of epidemiology of representations, do cultural attractors appear and how do cognitive biases participate in them?
We chose to restrict our analysis to substitutions (one word being replaced by another), both to keep the analysis tractable and because of missing information in our data set.\footnote{
\newtext{As explained further down, source-destination links between quotes must be inferred from the data set, an operation which is much more reliable if we restrict our analysis to substitutions.
This also impedes us from observing the effect of accumulated transformations in the long term, limiting our results to a view of the individual evolutionary step.
}}
While this limits the scope of our results to the particular data set we use, the methodological point we also make is left intact.
By characterizing words using 6 well-studied features, we identify what makes a substitution more likely, and how a word changes when it is substituted.
This exploratory approach uncovers a number of transmission biases consistent with known effects in linguistics.
While the transformations we describe are not the only ones at work in this data set, our analysis also indicates that feature-specific attractors could exist because of the substitution process.
This study can be viewed as analyzing part of the transmission step operating in transmission chains of artificial languages like those studied by \citet{kirby_cumulative_2008}, but with natural language out of the laboratory.

The next section %(Sec.~\ref{sec:related})
describes our hypotheses along with a review of the psycholinguistics literature.
%In Sec.~\ref{sec:protocol},
Then, we describe the data set and detail the various assumptions that were made in order to analyze it.
%Sec.~\ref{sec:results}
Next, we describe the measures built to observe the cognitive biases operating in quote transmission.
Finally, we discuss the relevance of these results for the study of cultural evolution, followed with general guidelines for further work.%Sec.~\ref{sec:conclusion}.

@@

The study of \newtext{cultural evolution on the part of cognitive science} emerged only recently.
\newtext{While formal models of cultural transmission appeared with the development of dual inheritance theory~\citep{cavalli-sforza_cultural_1981,boyd_culture_1985} and have included the notion of cultural attractor since then~\citep{claidiere_role_2007,claidiere_how_2014}, collecting data to test and iterate over such models has been more challenging.
The first method mentioned above consists in rebuilding the history of a given type of representation by compiling anthropological or historical works on the subject (as for instance \citealp{morin_how_2013}, and \citealp{miton_universal_2015}, have done).\footnote{
\newtext{Critics like \citet{ingold_trouble_2007}, however, have noted that the quantitative use of data imported from the social sciences risks overlooking the ontological debates in history and anthropology over the way to interpret such data.
}}
A second approach uses cultural evolution experiments in the laboratory, with an array of methods reviewed by \citet{mesoudi_multiple_2008}.
Transmission chains, in particular, have been used extensively to study the evolution of human language~\citep[see][for a review]{tamariz_cultural_2016}.
Other recent examples include studies of the evolution of simple audio loops through consumer preference~\citep{maccallum_evolution_2012}, the emergence of structure in visual patterns transmitted by baboons~\citep{claidiere_cultural_2014}, and the amplification of risk perception through chains of casual conversation~\citep{moussaid_amplification_2015}.}

@@

\citet{potter_regeneration_1990} suggest that immediate recall of sentences is based on the retention of an unordered list of words which is then regenerated as a sentence at the moment of production.
Priming recall with other words can lead to replacement in the recalled sentence if the primed words support the overall meaning of the sentence.
Regenerated syntax can also be influenced by priming recall with another syntactic structure~\citep{potter_syntactic_1998}, or with verbs whose category constraints call for a different structure~\citep{lombardi_regeneration_1992}.

>>>>

\tb{The reciprocal influence between cognition and culture has a long history in} both social science and psychology. While this \tb{question} has been the subject of intense debate in the social sciences in the 20th century\tbn{ (starting with Durkheim's initial works, \citeyear{durkheim_les_1912}, later tackled in earnest by \hbox{e.g.} Mauss' \emph{Techniques of the Body}, \citeyear{mauss_les_1936}, Giddens' \emph{Structuration Theory}, \citeyear{giddens_constitution_1984}, and Bourdieu's \emph{Sens Pratique}, \citeyear{bourdieu_sens_1980})}, today's discussion is mostly structured by proponents from cognitive science.
%Since the very beginnings of both social science and psychology, scientists have tried to capture the way cognition and culture influence each other. The question has been the subject of intense debate in the social sciences in the 20th century (starting with Durkheim's initial works, \citeyear{durkheim_les_1912}, later tackled in earnest by \hbox{e.g.} Mauss' \emph{Techniques of the Body}, \citeyear{mauss_les_1936}, Giddens' \emph{Structuration Theory}, \citeyear{giddens_constitution_1984}, and Bourdieu's \emph{Sens Pratique}, \citeyear{bourdieu_sens_1980}), but today's discussion is mostly structured by proponents from cognitive science.
These construe culture as an evolutionary process analogous and parallel to biological evolution \tbn{(and especially the modern synthesis' account of it)}.
That analogy can \tb{also} be traced a long way back \tbn{in the 20th century and earlier}, with milestones such as Kroeber's works~\citeyearpar{kroeber_nature_1952}, Dawkins' \emph{Memetics}~\tb{\citeyearpar{dawkins_selfish_1976}}, and later the development of \emph{Dual Inheritance Theory} by \citet{boyd_culture_1985} and \citet{cavalli-sforza_cultural_1981} among others.
More recently, Dan Sperber has drawn on this principle to explicitly connect anthropology and cognitive science through the theory of \emph{Epidemiology of Representations} \citep{sperber_explaining_1996}, and the study of cultural evolution has been growing steadily since.

The collection of works by \citet{aunger_darwinizing_2000} (in particular \citealp{bloch_well-disposed_2000}, and \citealp{kuper_if_2000}) has shown how \tbn{the theory of} memetics cannot account for the levels of transformation culture undergoes as it is transmitted.
\Citet{mesoudi_multiple_2008} \tbn{have} discussed the uses of transmission chain experiments to test what dual inheritance theory can explain about cultural evolution.
\Citet{morin_how_2013} and \citet{miton_universal_2015}, by carefully compiling a series of anthropological works, \tb{demonstrate} how cognitive biases have influenced the evolution of cultural artifacts over several centuries.
\citeauthor{kirby_cumulative_2008}~(\citeyear{kirby_cumulative_2008,cornish_systems_2013}) have shown how evolutionary pressures lead to the emergence of structured and expressive artificial languages in simulations and laboratory experiments.
Such transmission chain experiments have also been explored in non-human primates by \citet{claidiere_cultural_2014}.

The theory of epidemiology of representations proposes a unifying framework for all these works by recasting them as questions of spread and transformation of representations:
these are alternatively located in the mind (``mental representations'' in Sperber's terminology), or in the outer world (``public representations'') as expressions of mental representations in diverse cultural artifacts (pieces of text, utterances, pictures, building techniques, etc.).
A human society is then modeled as a large dynamical system of people constantly interpreting public representations into mental representations, and producing new public representations based on what they have previously interpreted.
Two key points are that (a) transmission is not reliable (representations change significantly each time they are interpreted and produced anew, as opposed to \hbox{e.g.} memetics), and (b) the reciprocal influences of cognition and culture can be captured by studying the evolution of public representations themselves, which is what the \tb{above cited} studies are doing.

The theory makes an additional strong hypothesis, which this paper focuses on:
as transformations accumulate, some representations evolve to be very stable and spread throughout an entire society without changing any more (they are called ``cultural representations'', because they characterize a given culture).
This process should manifest itself as attractors (called ``cultural attractors'') in the dynamical system that models cultural evolution, that is:
there should be areas of the representation space where cognitive effects in transformations bring representations closer to a given stable asymptotic point.\footnote{
\newtext{Attractors need not be points in fact, they can also be sub-areas; in that case any transformation brings representations in the area closer to (or \tb{confined to}) the target sub-area.
}}

This hypothesis, a cornerstone of the theory because of the intelligibility it gives to cultural evolution, has been hard to test in concrete situations as quantitative data on out-of-laboratory cultural artifacts is not easy to collect.
One approach, as mentioned above, has been the meta-analysis of large bodies of anthropological studies (see \citealp{miton_universal_2015}, for instance).
This paper exemplifies a second approach, taking advantage of the ever-increasing avalanche of available digital footprints since the 2000's.
Indeed, tools and computing power to analyze such data are now widespread, and the body of research aimed at describing online communities and content is growing accordingly.
For instance, the propagation of cultural artifacts across social networks has been studied in blogspace~\citep{gruhl_information_2004} and in \tb{emails}~\citep{liben-nowell_tracing_2008};
\citet{cointet_socio-semantic_2009} \tbn{have }described the reciprocal influence between the social network topology and the distribution of issues;
\citet{leskovec_meme-tracking_2009} detailed the characteristic times and diffusion cycles both within these social networks and with respect to the topical dynamics of news media, and \citet{danescu-niculescu-mizil_you_2012} \tbn{have }studied the characteristics of particularly memorable quotes that circulate in those networks.
We believe \tb{these} works can connect the field of cultural evolution with psycholinguistics to advance the testing of cultural attractors.

@@

To show this we analyze \tb{how} quotes in blogs and media outlets are modified when they are copied from website to website.
These public representations should normally not change as they spread on the Web (as opposed to more elaborate expressions or opinions, not identified as quoted utterances), but empirical observation shows that they are in fact occasionally transformed~\citep{simmons_memes_2011}:
authors spontaneously transform quotes, not only cropping them but also replacing words.
For instance the quote ``we will not be scared of these cowards'' (a substring of a quote from former Pakistani President Asif Ali Zardari) is also found as ``we will not be \textbf{afraid} of these cowards''.
More meaningful changes often happen too, such as the transformation of McCain's ``I admire Senator Obama and his accomplishments'' during the 2008 U.S. presidential campaign, into ``I \textbf{respect} Senator Obama and his accomplishments''.
Since authors are implicitly required to copy quotes exactly, we can assume that most transformations, especially simple ones such as those shown above, are the result of automatic (\hbox{i.e.} hard to control) low-level cognitive biases of the authors.

\tb{We thus ask the following question}: given such representations that seem to evolve precisely because of the kind of automatic cognitive biases \tb{evoked} in the theory of epidemiology of representations, %do cultural attractors appear and 
how do cognitive biases participate in \tb{the appearance of cultural attractors}?
We chose to restrict our analysis to substitutions (i.e., one word being replaced by another), both to keep the analysis tractable and because of missing information in our data set.\footnote{
\newtext{As explained further down, source-destination links between quotes must be inferred from the data set, an operation which is much more reliable if we restrict our analysis to substitutions.
This also impedes us from observing the effect of accumulated transformations in the long term, limiting our results to a view of the individual evolutionary step.
}}
While this limits the scope of our results to the particular data set we use, the methodological point we also make is left intact.  By characterizing words using 6 well-studied features, we identify what makes a substitution more likely, and how a word changes when it is substituted.
This exploratory approach uncovers a number of transmission biases consistent with known effects in linguistics.
While the transformations we describe are not the only ones at work in this data set, our analysis also indicates that feature-specific attractors could exist because of the substitution process.
This study can be viewed as analyzing part of the transmission step operating in transmission chains of artificial languages like those studied by \citet{kirby_cumulative_2008}, \tb{yet} with natural language out of the laboratory.

The next section %(Sec.~\ref{sec:related})
describes our hypotheses along with a review of the psycholinguistics literature.
%In Sec.~\ref{sec:protocol},
Then, we describe the data set and detail the various assumptions that were made in order to analyze it.
%Sec.~\ref{sec:results}
Next, we \tb{introduce} the measures \tb{we} built to observe \tbn{the} cognitive biases operating in quote transmission. Finally, we discuss the relevance of these results for the study of cultural evolution, followed with general guidelines for further work.%Sec.~\ref{sec:conclusion}.
\end{new}

@@

The study of \newtext{cultural evolution on the part of cognitive science} emerged only recently.
\newtext{While formal models of cultural transmission appeared with the development of dual inheritance theory~\citep{cavalli-sforza_cultural_1981,boyd_culture_1985} and have included the notion of cultural attractor since then~\citep{claidiere_role_2007,claidiere_how_2014}, collecting data to test and iterate over such models has been more challenging.
The first \tb{above-mentioned} method consists in rebuilding the history of a given type of representation by compiling anthropological or historical works on the subject (as for instance \citealp{morin_how_2013}, and \citealp{miton_universal_2015}, have done).\footnote{\tbn{Critics like }\tb{\citet{ingold_trouble_2007}}\tbn{, however, have noted that the quantitative use of data imported from the social sciences risks overlooking the ontological debates in history and anthropology over the way to interpret such data.}}
A second approach uses cultural evolution experiments in the laboratory, with an array of methods reviewed by \citet{mesoudi_multiple_2008}.
Transmission chains, in particular, have been used extensively to study the evolution of human language~\citep[see][for a review]{tamariz_cultural_2016}.
Other recent examples include studies of the evolution of simple audio loops through consumer preference~\citep{maccallum_evolution_2012}, the emergence of structure in visual patterns transmitted by baboons~\citep{claidiere_cultural_2014}, and the amplification of risk perception through chains of casual conversation~\citep{moussaid_amplification_2015}.}

@@

\citet{potter_regeneration_1990} suggest that immediate recall of sentences is based on the retention of an unordered list of words which is then regenerated as a sentence at the moment of production.
Priming recall with other words can lead to replacement in the recalled sentence if the primed words \tb{are consistent with} the overall meaning of the sentence.
Regenerated syntax can also be influenced by priming recall with another syntactic structure~\citep{potter_syntactic_1998}, or with verbs whose category constraints call for a different structure~\citep{lombardi_regeneration_1992}.
